Effects of sexual selection on non-sexual fitness components: a meta-analysis

### 2018-02-27 Working Draft/Meta-Analysis on Sexual Selection and Population Fitness.Rmd
Add files via upload

```{r}
prelim.data$Study.ID <- prelim.data$Study.ID %>% factor
prelim.data$Taxon <- prelim.data$Taxon %>% factor
prelim.data$Experiment.ID <- prelim.data$Experiment.ID %>% factor 
prelim.data$Authors <- prelim.data$Authors %>% factor
prelim.data$Environment <- prelim.data$Environment %>% factor %>% relevel(ref="Unstressed")
prelim.data$Sex <- prelim.data$Sex %>% factor %>% relevel(ref="B")
prelim.data$Ambiguous <- prelim.data$Ambiguous %>% factor
prelim.data$Species <- prelim.data$Species %>% factor
#Outcome.Class.2 is using the categories that were decided by survey. I am keeping both just to check them against each other (how much of a difference it makes)
prelim.data$Outcome.Class <- prelim.data$Outcome.Class %>% factor %>% relevel(ref="Indirect")
prelim.data$Enforced.Monogamy <- prelim.data$Enforced.Monogamy %>% factor %>% relevel(ref="NO")
prelim.data$Pre.cop <- prelim.data$Pre.cop %>% factor %>% relevel(ref="0")
prelim.data$Post.cop <- prelim.data$Post.cop %>% factor %>% relevel(ref="0")
prelim.data$Ratio.Category <- prelim.data$Ratio.Category %>% factor %>% relevel(ref="Low")
prelim.data$Density.Category <- prelim.data$Density.Category %>% factor %>% relevel(ref="Low")
prelim.data$SSS.Categorical <- prelim.data$SSS.Categorical %>% factor %>% relevel(ref="Low")
prelim.data$Blinding <- prelim.data$Blinding %>% factor
```

@@

```{r}
n.blind.ones <- (sum(prelim.data$Blind == "Blind"))
n.stressed.ones <- 
prelim.data %>% summarise(Effect_sizes = n(), 
                       Publications = prelim.data$Study.ID %>% unique %>% length,
                       Blind_experiments = n.blind.ones,
                       Effect_sizes_.Stressedq = (sum(prelim.data$Environment == "Stressed")),
                       Effect_sizes_.Unstressedq = (sum(prelim.data$Environment == "Unstressed")),         
                       Effect_sizes_.Maleq = (sum(prelim.data$Sex == "M")),
                       Effect_sizes_.Femaleq = (sum(prelim.data$Sex == "F")),
                       Different_species =  prelim.data$Species %>% unique %>% length,
                       Effect_sizes_.Beetleq = sum(Taxon == "Beetle"),
                       Effect_sizes_.Flyq = sum(Taxon == "Fly"),
                       Effect_sizes_.Mouseq = sum(Taxon == "Mouse"),
                       Effect_sizes_.Nematodeq = sum(Taxon == "Nematode"),
                       Effect_sizes_.Miteq = sum(Taxon == "Mite"),
                       Effect_sizes_.Cricketq = sum(Taxon == "Cricket"),
                       Effect_sizes_.Guppyq = sum(Taxon == "Guppy")) %>% melt %>%
  mutate(variable = gsub("_", " ", variable),
         variable = gsub("[.]", "(", variable),
         variable = gsub("q", ")", variable)) %>% 
  rename_("n" = "value", " " = "variable") %>% pander(split.cell = 40, split.table = Inf)
```

@@

Here we show the residual effect sizes for each outcome measured. The following forest plot is based on the residual effect sizes. The model is an intercept only model with the same multilevel random effects structure as we use later when investigating environment and sex

@@

```{r, fig.height=30, fig.width=22.5}
#Create model we later establish and analyse but not restricted
forest.model <- rma.mv(g, var.g, mods = ~ 1, random = ~ 1 | Study.ID/Outcome, method = "REML", data = prelim.data)

#Obtain residuals
resstandards <- (rstandard.rma.mv(forest.model, type="response"))

@@

pd <- position_dodgev(height = .3)
predictions.complete2 %>% ggplot(aes(x = pred, y= Environment, colour = Sex)) + 
  geom_vline(xintercept = 0, linetype = 2, colour = "grey70") + 
  geom_errorbarh(aes(xmin = predictions.complete2$ci.lb, xmax = predictions.complete2$ci.ub), height = 0, position = pd) +
  geom_point(position = pd) + 
  ylab("Environment")+
  xlab("Effect Size (Hedges g)")+
  xlim(-.75, .75)+
  ggtitle('Effects of Sex, Stress \non Population Fitness')
```

@@

> This inner, outer factor stuff from the metafor package is a bit strange. There is a goodish description in ``?rma.mv()`` but still unsure how it differs from the above model. It seems that it is useful to breakdown variance-covariance matrix but unsure how that would benefit our analysis.

@@

# For stressed environment and females
varSF <- rma.mv(lnCVr, var.lnCVr, data = prelim.data2, mods = ~ 1, random = ~ 1 | Study.ID, subset = (Environment == "Stressed" & Sex == "F"))

# For stressed environment and females
varSM <- rma.mv(lnCVr, var.lnCVr, data = prelim.data2, mods = ~ 1, random = ~ 1 | Study.ID, subset = (Environment == "Stressed" & Sex == "M"))

# For stressed environment and females
varSB <- rma.mv(lnCVr, var.lnCVr, data = prelim.data2, mods = ~ 1, random = ~ 1 | Study.ID, subset = (Environment == "Stressed" & Sex == "B"))

#For Benign environment and females
varUF<- rma.mv(lnCVr, var.lnCVr, data = prelim.data2, mods = ~1, random = ~ 1 | Study.ID, subset = (Environment == "Unstressed" & Sex == "F"))

#For Benign environment and males
varUM<- rma.mv(lnCVr, var.lnCVr, data = prelim.data2, mods = ~1, random = ~ 1 | Study.ID, subset = (Environment == "Unstressed" & Sex == "M"))

#For Benign environment and both
varUB <- rma.mv(lnCVr, var.lnCVr, data = prelim.data2, mods = ~1, random = ~ 1 | Study.ID, subset = (Environment == "Unstressed" & Sex == "B"))

@@

Checking for biases with a funnell plot. Note that the trim and fill method does not work with rma.mv objects. However we can perform Eggers test using the ``ragtest()`` function. This tests for asymmetry via assessing relationships between effect size and a specified predictor. See ``?ragtest()`` for more information. Also the eggers test does not work for rma.mv objects. 
```{r}
funnel.rma(model.Ambiguous)
regtest(model.Ambiguous, predictor = "vi")
funnel.rma(model.Indirect)
regtest(model.Indirect)
funnel.rma(model.Direct)
regtest(model.Direct)
funnel.rma(model.complete2)
```

@@

If we see a positive trend with effect size and Journal Impact Factor it may represent publication bias whereby significant (positive) results are published more readily and in more circulated journals. Our journal impact factor dataset is not evenly distributed as several publications in Nature (JIF ~ 40) are much larger than the next highest JIF (~11)

```{r}
prelim.data %>% ggplot(aes(x=JIF, y=(g), size = 1/sqrt(var.g)))+
  geom_jitter()+
  geom_hline(yintercept=0, linetype = 'dotted')+
  geom_smooth(method = 'lm')
```

@@

@@

```{r}
prelim.data %>% ggplot(aes(x=Year, y=g, size = (1/(var.g))))+
  geom_jitter()+
  geom_hline(yintercept=0, linetype = 'dotted')+
  geom_smooth(method='lm')+
  labs(size = 'weight')
```

@@

```{r}
df.forest.model %>% ggplot(aes(x=Blinding, y=g, color = Outcome.Class))+
  geom_boxplot()+
  geom_hline(yintercept=0, linetype = 'dotted')
```

>>>>

```{r}
prelim.data$Study.ID <- prelim.data$Study.ID %>% factor
prelim.data$Taxon <- prelim.data$Taxon %>% factor
prelim.data$Group.ID <- prelim.data$Group.ID %>% factor 
prelim.data$Authors <- prelim.data$Authors %>% factor
prelim.data$Environment <- prelim.data$Environment %>% factor %>% relevel(ref="Unstressed")
prelim.data$Sex <- prelim.data$Sex %>% factor %>% relevel(ref="B")
prelim.data$Ambiguous <- prelim.data$Ambiguous %>% factor
prelim.data$Species <- prelim.data$Species %>% factor
#Outcome.Class.2 is using the categories that were decided by survey. I am keeping both just to check them against each other (how much of a difference it makes)
prelim.data$Outcome.Class <- prelim.data$Outcome.Class %>% factor %>% relevel(ref="Indirect")
prelim.data$Enforced.Monogamy <- prelim.data$Enforced.Monogamy %>% factor %>% relevel(ref="NO")
prelim.data$Pre.cop <- prelim.data$Pre.cop %>% factor %>% relevel(ref="0")
prelim.data$Post.cop <- prelim.data$Post.cop %>% factor %>% relevel(ref="0")
prelim.data$Ratio.Category <- prelim.data$Ratio.Category %>% factor %>% relevel(ref="Low")
prelim.data$Density.Category <- prelim.data$Density.Category %>% factor %>% relevel(ref="Low")
prelim.data$SSS.Categorical <- prelim.data$SSS.Categorical %>% factor %>% relevel(ref="Low")
prelim.data$Blinding <- prelim.data$Blinding %>% factor
```

@@

```{r}
n.blind.ones <- (sum(prelim.data$Blind == "Blind"))
n.stressed.ones <- 
prelim.data %>% summarise(Effect_sizes = n(), 
                       Publications = prelim.data$Study.ID %>% unique %>% length,
                       Blind_experiments = n.blind.ones,
                       Effect_sizes_.Stressedq = (sum(prelim.data$Environment == "Stressed")),
                       Effect_sizes_.Unstressedq = (sum(prelim.data$Environment == "Unstressed")),         
                       Effect_sizes_.Maleq = (sum(prelim.data$Sex == "M")),
                       Effect_sizes_.Femaleq = (sum(prelim.data$Sex == "F")),
                       Effect_sizes_.Both_sexesq = (sum(prelim.data$Sex == "B")),
                       Different_species =  prelim.data$Species %>% unique %>% length,
                       Effect_sizes_.Beetleq = sum(Taxon == "Beetle"),
                       Effect_sizes_.Flyq = sum(Taxon == "Fly"),
                       Effect_sizes_.Mouseq = sum(Taxon == "Mouse"),
                       Effect_sizes_.Nematodeq = sum(Taxon == "Nematode"),
                       Effect_sizes_.Miteq = sum(Taxon == "Mite"),
                       Effect_sizes_.Cricketq = sum(Taxon == "Cricket"),
                       Effect_sizes_.Guppyq = sum(Taxon == "Guppy")) %>% melt %>%
  mutate(variable = gsub("_", " ", variable),
         variable = gsub("[.]", "(", variable),
         variable = gsub("q", ")", variable)) %>% 
  rename_("n" = "value", " " = "variable") %>% pander(split.cell = 40, split.table = Inf)
```

@@

Here we show the residual effect sizes for each outcome measured. The following forest plot is based on the residual effect sizes. The model is an intercept only model with standard random effects structure.

@@

```{r, fig.height=30, fig.width=22.5}
#Create standard random effectrs model
forest.model <- rma(g, var.g, mods = ~ 1, method = "REML", data = prelim.data)

#Obtain residuals
resstandards <- (rstandard.rma.uni(forest.model, type="response"))

@@

pd <- position_dodgev(height = .5)
predictions.complete2 %>% ggplot(aes(x = pred, y= Environment, colour = Sex)) + 
  geom_vline(xintercept = 0, linetype = 2, colour = "grey70") + 
  geom_errorbarh(aes(xmin = predictions.complete2$ci.lb, xmax = predictions.complete2$ci.ub), height = 0, position = pd) +
  geom_point(position = pd) + 
  ylab("Environment")+
  xlab("Effect Size (Hedges g)")+
  xlim(-.75, .75)+
  ggtitle('Effects of Sex, Stress \non Population Fitness')
```

We see that female fitness in stressed environments is greater than the other measurements. For outcomes that were measured for both female and males we see a greater uncertainty in the estimate. It is not obviously clear why this is. The 'both' outcomes are restricted to extinction rate, offspring viability, mutant frequency and reproductive success. However, the shift from 'both' being ns in unstressed to significant in stressed may reflect the dampening of the negative correlations (sexual antagonism).

```{r}
table(restricted.data$Outcome, restricted.data$Sex)
```

@@

> This inner, outer factor stuff from the metafor package is a bit strange. There is a description in ``?rma.mv()`` but still unsure how it differs from the above model. It seems that it is useful to breakdown variance-covariance matrix but unsure how that would benefit our analysis.

@@

# For stressed environment and females
varSF <- rma.mv(lnCVr, var.lnCVr, data = prelim.data2, mods = ~ 1, random = ~ 1 | Study.ID/Outcome, subset = (Environment == "Stressed" & Sex == "F"))

# For stressed environment and females
varSM <- rma.mv(lnCVr, var.lnCVr, data = prelim.data2, mods = ~ 1, random = ~ 1 | Study.ID/Outcome, subset = (Environment == "Stressed" & Sex == "M"))

# For stressed environment and females
varSB <- rma.mv(lnCVr, var.lnCVr, data = prelim.data2, mods = ~ 1, random = ~ 1 | Study.ID/Outcome, subset = (Environment == "Stressed" & Sex == "B"))

#For Benign environment and females
varUF<- rma.mv(lnCVr, var.lnCVr, data = prelim.data2, mods = ~1, random = ~ 1 | Study.ID/Outcome, subset = (Environment == "Unstressed" & Sex == "F"))

#For Benign environment and males
varUM<- rma.mv(lnCVr, var.lnCVr, data = prelim.data2, mods = ~1, random = ~ 1 | Study.ID/Outcome, subset = (Environment == "Unstressed" & Sex == "M"))

#For Benign environment and both
varUB <- rma.mv(lnCVr, var.lnCVr, data = prelim.data2, mods = ~1, random = ~ 1 | Study.ID/Outcome, subset = (Environment == "Unstressed" & Sex == "B"))

@@

Checking for biases with a funnell plot. Note that the trim and fill method does not work with rma.mv objects. However we can perform Eggers test using the ``ragtest()`` function. This tests for asymmetry via assessing relationships between effect size and a specified predictor. See ``?ragtest()`` for more information. Also the eggers test does not work for rma.mv objects. 
```{r}
trim.ambig <- trimfill.rma.uni(model.Ambiguous)
funnel.rma(trim.ambig)
regtest(model.Ambiguous, predictor = "vi")
funnel.rma(model.Indirect)
regtest(model.Indirect, predictor = "vi")
funnel.rma(model.Direct)
regtest(model.Direct, predictor = "vi")
funnel.rma(model.complete2, type="rstandard", yaxis = "sei")
```

@@

####Journal Impact Factor

If we see a positive trend with effect size and Journal Impact Factor it may represent publication bias whereby significant (positive) results are published more readily and in more circulated journals. Our journal impact factor dataset is not evenly distributed as several publications in Nature (JIF ~ 40) are much larger than the next highest JIF (~11).

```{r}

```{r, fig.width=10, fig.height=8}
prelim.data %>% ggplot(aes(x=JIF, y=(g)))+
  geom_jitter(color='darkgreen', alpha=.3, aes(size = (1/(var.g))))+
  geom_hline(yintercept=0, linetype = 'dotted')+
  geom_smooth(method='lm', color='black')+
  labs(size = 'weight')
```

@@

####Time-lag Bias

@@

```{r, fig.width=7.50, fig.height=6}
prelim.data %>% ggplot(aes(x=Year, y=g))+
  geom_jitter(color='orange', alpha=.3, aes(size = (1/(var.g))))+
  geom_hline(yintercept=0, linetype = 'dotted')+
  geom_smooth(method='lm', color='black')+
  labs(size = 'weight')
```

####Blinding

@@

```{r, fig.width=10, fig.height=8}
df.forest.model %>% ggplot(aes(x=Blinding, y=g))+
  geom_boxplot()+
  geom_jitter(aes(color=Blinding))+
  geom_hline(yintercept=0, linetype = 'dotted')
```

####Sample Size

We also collected sample sizes for each of the effect sizes calculated. Because we are dealing withj different taxa some studies are not suited to have sample sizes in the 1000's. We can simply inspect the sample size ande effect sizes through the following plot:
```{r, fig.width=5, fig.height=10}
prelim.data %>% ggplot(aes(x=(n), y = g))+
  geom_point()+
  xlim(0,2100)+
  facet_grid(Taxon~., scales='free')+
  ylim(-3.5,3.5)+
  geom_hline(yintercept=0, linetype="dashed")
```

(Promislow 1998 has one sample size of >10,000) and is not shown here. From these plots we can see that with increased sample size the effect sizes are closer to zero. Thistrend should be taken into account as meta-analytic models are wighted by 1/variance.

####Generations

We recorded the number of generations of experimental exolution each study used. The number of generations proved a negligable predictor in the meta-analytic models and can be seen here:

```{r, fig.width=8, fig.height=7}
prelim.data %>% ggplot(aes(x=Generations, y=g, color=Taxon))+
  geom_point()+
  ylim(-3.5,3.5)+
  geom_hline(yintercept=0, linetype="dashed")
```
